---
title: "Untitled"
output: html_document
date: '2022-05-06'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Packages
```{r packages, include=FALSE}

library(dplyr)
library(Seurat)
library(ggplot2)
library(colorspace)
library(STACAS)
library(clustree)
library(gridExtra)
library(ggpubr)
library(egg)
library(scales)
library(scran)
library(effectsize)

```

###--------------- DATA LOADING -------------------- ###
```{r read.data, include=FALSE}
counts <- read.csv('~/Desktop/SCRNA/datasets/SCP503/other/Richards_NatureCancer_GBM_scRNAseq_counts.csv',
                   header = TRUE,
                   row.names = 1,
                   sep = ",")

colnames(counts) <- gsub(x = colnames(counts),
                         pattern = "\\.",
                         replacement = "-")

meta <- read.csv('~/Desktop/SCRNA/datasets/SCP503/other/Richards_NatureCancer_GBM_scRNAseq_meta.csv',
                 header = TRUE,
                 row.names = 1,
                 sep = ",")

all(rownames(meta)==colnames(counts))

```

### SEURAT PIPELINE ###
```{r seurat.obj, include=TRUE}
GBM <- CreateSeuratObject(counts = counts,
                          project = "GBM",
                          meta.data = meta)

GBM[["ribo"]] <- PercentageFeatureSet(GBM, pattern = "^RP[SL]")
GBM[["mito"]] <- PercentageFeatureSet(GBM, pattern = "^MT-")

saveRDS(GBM, "~/Desktop/SCRNA/outputs/GBM.rds")
```
### SAMPLE LISTS ###
```{r }
#red_list <- list()
GBM <- readRDS("~/Desktop/SCRNA/outputs/GBM.rds")
red_list <- c( 'G1003-A_T', 'G1003-B_T', 'G1003-C_T')

GBMr <- subset(GBM, subset=orig.ident %in% red_list)
rm(GBM)
sample.list <- SplitObject(GBMr, split.by = "orig.ident")
sce.list <- lapply(sample.list, as.SingleCellExperiment)

```

### PLOTTING ###
```{r plots.QC, fig.height=4, fig.width=9}
ggplot(GBMr[[]], aes(nCount_RNA, mito, color=ribo)) +
  geom_jitter(size=0.3) +
  theme_pubr() +
  #scale_y_log10() +
  scale_x_log10() +
  scale_color_continuous_sequential(palette='ag_Sunset', rev=T) 
 # facet_wrap(~orig.ident, nrow=2)

ggplot(GBMr[[]], aes(nGene, mito, color=PatientID)) +
  geom_jitter(size=0.3) +
  theme_pubr() +
  #scale_y_log10() +
  scale_x_log10() +
  #scale_color_continuous_sequential(palette='ag_Sunset', rev=T) +
  facet_wrap(~orig.ident, nrow=2)

ggplot(GBMr[[]], aes(nCount_RNA, nGene, color=ribo)) +
  geom_jitter(size=0.3) +
  theme_pubr() +
  scale_y_log10() +
  scale_x_log10() +
  scale_color_continuous_sequential(palette='ag_Sunset', rev=T)

ggplot(GBMr[[]], aes(nCount_RNA, nGene, color=mito)) +
  geom_jitter(size=0.3) +
  theme_pubr() +
  scale_y_log10() +
  scale_x_log10() +
  scale_color_continuous_sequential(palette='ag_Sunset', rev=T)

ggplot(GBMr[[]], aes(nCount_RNA, nGene, color=orig.ident)) +
  geom_jitter(size=0.3) +
  theme_pubr() +
  scale_y_log10() +
  scale_x_log10() 

ggplot(GBMr[[]], aes(nCount_RNA, nGene, color=nGene<200)) +
  geom_jitter(size=0.3) +
  theme_pubr() +
  scale_y_log10() +
  scale_x_log10() 

```

### INITIAL INTEGRATION ###
``` {r integration, include=TRUE}

for (i in 1:length(sample.list)) {
  sample.list[[i]] <- SCTransform(sample.list[[i]])
  #sample.list[[i]] <- NormalizeData(sample.list[[i]]) %>% FindVariableFeatures(., selection.method = 'vst', nfeatures=3000) %>% ScaleData(.)
  #clust <- quickCluster(sce.list[[i]])
  #sce <- computeSumFactors(sce.list[[i]], cluster=clust, min.mean=0.1)
  #sce <- scater::logNormCounts(sce)
  #sample.list[[i]] <- CreateAssayObject(data=sce@assays@data$logcounts)
}
```

### MERGING ###
```{r, include=F }

m <- sample.list[[1]]
for (i in 2:length(sample.list)){
  m <- merge(m, sample.list[[i]])
}
```

### CLUSTERING ###
```{r, include=F }

m <- SCTransform(m)
m <- RunPCA (m)
m <- RunUMAP(m, 
            dims=1:30,
            n.neighbors = 10)

m <- FindNeighbors(m, 
                  reduction ="pca")

m  <- FindClusters(m)

```
### PLOT CLUSTERS ###
```{r }

DimPlot(m, 
        group.by='SampleID')
FeaturePlot(m, features=c('PTPRC', 'CD24', 'FAM107A', 'HLA-A'), order=T)
DimPlot(m, 
        group.by='seurat_clusters', label=T)


```

### subsettting out immune cells and cancr cells
```{r }
`%ni%` <- Negate(`%in%`)
immune <- subset(m, subset=seurat_clusters  %in% c(0,7))
tumor <- subset(m, subset=seurat_clusters  %ni% c(0,7))
FeaturePlot(immune, features=c('PTPRC', 'CD24', 'FAM107A', 'HLA-A'), order=T)
FeaturePlot(tumor, features=c('PTPRC', 'CD24', 'FAM107A', 'HLA-A'), order=T)
DimPlot(immune, 
        group.by='seurat_clusters', label=T)
DimPlot(tumor, 
        group.by='seurat_clusters', label=T)

```
### split sample/sce list into tumour & immune cells ### 
```{r split, include=F}
sample.list.immune <- SplitObject(immune, split.by = "orig.ident")
sce.list.immune <- lapply(sample.list.immune, as.SingleCellExperiment)

sample.list.tumor <- SplitObject(tumor, split.by = "orig.ident")
sce.list.tumor <- lapply(sample.list.tumor, as.SingleCellExperiment)

```

### transform again - immune & tumour cells ###
```{r transform, include=F}
for (i in 1:length(sample.list.immune)) {
  sample.list.immune[[i]] <- SCTransform(sample.list.immune[[i]])
  #sample.list[[i]] <- NormalizeData(sample.list[[i]]) %>% FindVariableFeatures(., selection.method = 'vst', nfeatures=3000) %>% ScaleData(.)
  #clust <- quickCluster(sce.list[[i]])
  #sce <- computeSumFactors(sce.list[[i]], cluster=clust, min.mean=0.1)
  #sce <- scater::logNormCounts(sce)
  #sample.list[[i]] <- CreateAssayObject(data=sce@assays@data$logcounts)
}
for (i in 1:length(sample.list.tumor)) {
  sample.list.tumor[[i]] <- SCTransform(sample.list.tumor[[i]])
}
```

### merge tumour & immune ### 
```{r merge, include=F }

m.immune <- sample.list.immune[[1]]
for (i in 2:length(sample.list.immune)){
  m.immune <- merge(m, sample.list.immune[[i]])
}

m.tumor <- sample.list.tumor[[1]]
for (i in 2:length(sample.list.tumor)){
  m.tumor <- merge(m, sample.list.tumor[[i]])
}
```

### transform merged lists ###
```{r, include=F }

m.immune <- SCTransform(m.immune)
m.tumor <- SCTransform(m.tumor)

m.immune <- RunPCA (m.immune)
m.tumor <- RunPCA(m.tumor)

m.immune <- RunUMAP(m.immune, 
            dims=1:30,
            n.neighbors = 10)
m.tumor <- RunUMAP(m.tumor, 
            dims=1:30,
            n.neighbors = 10)

m.immune <- FindNeighbors(m.immune, 
                  reduction ="pca")
m.tumor <- FindNeighbors(m.tumor, 
                  reduction ="pca")

m.immune <- FindClusters(m.immune)
m.tumor <- FindClusters(m.tumor)

```

### Plotting reclustered immune & tumor###
```{r plot}

DimPlot(m.immune, 
        group.by='SampleID')
FeaturePlot(m.immune, features=c('PTPRC', 'CD24', 'ITGAM', 'HLA-A'), order=T)
DimPlot(m.immune, 
        group.by='seurat_clusters', label=T)

DimPlot(m.tumor, 
        group.by='SampleID')
FeaturePlot(m.tumor, features=c('PTPRC', 'CD24', 'ITGAM', 'HLA-A'), order=T)
DimPlot(m.tumor, 
        group.by='seurat_clusters', label=T)


```
### RE-SUBSET CELLS ###
```{r }
`%ni%` <- Negate(`%in%`)
immune1 <- subset(m.immune, subset=seurat_clusters  %ni% c(6)) # everything except cluster 6 from prev. clustered immune cells
immune2 <- subset(m.tumor, subset=seurat_clusters %in% c(0,10)) # cluster 0 & 10 from prev. clustered tumor cells
#immune <- merge(immune1, immune2)

tumor1 <- subset(m.immune, subset=seurat_clusters  %in% c(6)) # cluster 6 from prev. clustered immune cells
tumor2 <- subset(m.tumor, subset=seurat_clusters %ni% c(0, 10)) # everything except cluster 0 & 10 from prev. clustered tumor cells
#tumor <- merge(tumor1, tumor2)


FeaturePlot(immune1, features=c('PTPRC', 'CD24', 'FAM107A', 'HLA-A'), order=T)
FeaturePlot(immune2, features=c('PTPRC', 'CD24', 'FAM107A', 'HLA-A'), order=T)
FeaturePlot(tumor1, features=c('PTPRC', 'CD24', 'FAM107A', 'HLA-A'), order=T)
FeaturePlot(tumor2, features=c('PTPRC', 'CD24', 'FAM107A', 'HLA-A'), order=T)
DimPlot(immune1, 
        group.by='seurat_clusters', label=T)
DimPlot(immune2, 
        group.by='seurat_clusters', label=T)
DimPlot(tumor1, 
        group.by='seurat_clusters', label=T)
DimPlot(tumor2, 
        group.by='seurat_clusters', label=T)

```
### MERGE ###
```{r merge. include=F}
m.immune <- merge(immune1, immune2)
m.tumor <- merge(tumor1, tumor2)
```

### transform merged lists ###
```{r, include=F }

m.immune <- SCTransform(m.immune)
m.tumor <- SCTransform(m.tumor)

m.immune <- RunPCA (m.immune)
m.tumor <- RunPCA(m.tumor)

m.immune <- RunUMAP(m.immune, 
            dims=1:30,
            n.neighbors = 10)
m.tumor <- RunUMAP(m.tumor, 
            dims=1:30,
            n.neighbors = 10)

m.immune <- FindNeighbors(m.immune, 
                  reduction ="pca")
m.tumor <- FindNeighbors(m.tumor, 
                  reduction ="pca")
```
### cluster ###
```{r cluster}
m.immune <- FindClusters(m.immune, resolution = 0.2)
m.tumor <- FindClusters(m.tumor, resolution = 0.8)

```

### Plotting reclustered immune & tumor###
```{r plot}

DimPlot(m.immune, 
        group.by='SampleID')
FeaturePlot(m.immune, features=c('PTPRC', 'CD24', 'FAM107A', 'HLA-A', 'PECAM1'), order=T)
DimPlot(m.immune, 
        group.by='seurat_clusters', label=T)

DimPlot(m.tumor, 
        group.by='SampleID')
FeaturePlot(m.tumor, features=c('PTPRC', 'CD24', 'FAM107A', 'HLA-A', 'PECAM1'), order=T)
DimPlot(m.tumor, 
        group.by='seurat_clusters', label=T)
```
### Continue with STACAS integration

```{r STACAS, fig.width=24, fig.height=12, include=F}

features <- SelectIntegrationFeatures(object.list=sample.list, 
                                      nfeatures=900)

sample.list <- PrepSCTIntegration(object.list=sample.list, 
                                  anchor.features = features)

stacas.anchors <- FindAnchors.STACAS(sample.list, 
                                     anchor.features =features, 
                                     normalization.method = "SCT")

names<- names(sample.list)
plots <- PlotAnchors.STACAS(stacas.anchors, obj.names = names)
```

```{r plot, fig.width = 24, fig.height=12}
g <- do.call("arrangeGrob", c(plots))
plot(g)

stacas.anchors.filtered <- FilterAnchors.STACAS(stacas.anchors, dist.thr = 0.75)
```



```{r }
mySampleTree <- SampleTree.STACAS(stacas.anchors.filtered)

stacas.int <- IntegrateData(anchorset = stacas.anchors.filtered, 
                            k.weight = 14, 
                            normalization.method = "SCT",
                            dims = 1:10, 
                            sample.tree = mySampleTree, 
                            preserve.order = T)

```

``` {r pc.analysis, include=TRUE}

## Reductions - PCA, UMAP, tSNE
DefaultAssay(stacas.int) <- 'integrated'
stacas.int <- RunPCA(stacas.int, verbose=FALSE)

# Determine percent of variation associated with each PC
pct <- stacas.int[["pca"]]@stdev / sum(stacas.int[["pca"]]@stdev) * 100

# Calculate cumulative percents for each PC
cumu <- cumsum(pct)

# Determine which PC exhibits cumulative percent greater than 90% and % variation associated with the PC as less than 5
co1 <- which(cumu > 90 & pct < 5)[1]

# Determine the difference between variation of PC and subsequent PC

co2 <- sort(which((pct[1:length(pct) - 1] - pct[2:length(pct)]) > 0.1), decreasing = T)[1] + 1

stacas.int <- RunUMAP(stacas.int, 
                      dims=1:co2,
                      n.neighbors = 10,
                      spread = 0.3,
                      min.dist=0.001)

```

``` {r cluster, include=TRUE, fig.height=12, fig.width=24}

### Clustering using seurat
stacas.int <- FindNeighbors(stacas.int, 
                            reduction ="pca", 
                            dims=1:co2)

for (r in seq(0,0.9,0.1)){
  stacas.int  <- FindClusters(stacas.int , resolution=r)
}

#Plot clustree
clustree::clustree(stacas.int, prefix="integrated_snn_res.") + 
  ggtitle("Clustering tree (data.int)")

plots <- list()
for (res in seq(0,0.9, 0.1)) 
  plots[[as.character(res)]] <- DimPlot(stacas.int, pt.size = 0.5,
                                        label=T, repel=T,
                                        group.by = paste0("integrated_snn_res.", res)) + 
  ggtitle(paste("res=", res)) + theme(legend.position = "none")

cowplot::plot_grid(plotlist=plots)

#saveRDS(stacas.int, "~/Desktop/SCRNA/outputs/stacas.int_SCT_14scSamples.rds")

```
```{r, fig.width=12, fig.height=6}

DefaultAssay(stacas.int) <- 'SCT'
DimPlot(stacas.int, reduction="umap", group.by = "integrated_snn_res.0.5")
FeaturePlot(stacas.int, features=c('PTPRC','CD3','CD4', 'CD8A', 'CD11B', 'CD68', 'CD44', 'CD24'), order=T)
```
